Speech Enhancement for Low Quality Audio

### 2020-04-25 2020-tsd-brno-denoise/denoise-anonymous.tex
2020 denoise: use correct template and add anonymized version

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{rotating}

\usepackage{url}
\urldef{\mailsa}\path|kruza@ufal.mff.cuni.cz|    
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Mitigating Low Audio Quality by Domain Transfer for Better ASR}

% a short form should be given in case it is too long for the running head
\titlerunning{Audio Domain Transfer for Better ASR}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Jan Oldřich Krůza \and David Klusáček}
%
\authorrunning{Jan Oldřich Krůza}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Institute of Formal and Applied Linguistics,\\
Faculty of Mathematics and Physics,\\
Charles University
\mailsa\\
\url{ufal.mff.cuni.cz}}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

\toctitle{Mitigating Low Audio Quality by Domain Transfer for Better ASR}
\tocauthor{Jan Oldřich Krůza}
\maketitle

\begin{abstract}
The paper presents an attempt to solve the problem of bad audio quality
affecting the accuracy of speech-to-text systems. We
describe an experiment with Gycle-GAN domain transfer from the defect to the
pure as a way to reduce the impact of noise, overdrive, reverberation and other
defects.
\keywords{speech recognition, GycleGAN, domain transfer, signal processing}
\end{abstract}

@@

There are in principle two ways to deal with the negative effect of acoustic
deficiencies in speech recognition: 1) to adapt the model, which includes
training on augmented data, and 2) to adapt the input data. This article
presents an experiment in the latter approach.

@@

The application is at the spoken corpus of Karel Makoň\cite{makondata}, for which
a speech recognition system has been developed\cite{kruuza2012making} but where
the heterogenous acoustic quality affects both the speech recognition
performance and the listening experience. Whereas the word error rate on the
overall test set is 21\%, it is 45\% on a sample of recordings suffering from
overdrive, and 68\% on recordings taken on a magnetophone tape with a slow
recording speed of 2 centimeters per second after some 30 years of living-room
storage.

@@

The individual interference types affect each other and can occur several times
in several stages. For instance, the echo is essentially convolution with a certain vector of
amplitude values, known as the impulse response. It even might not be
constant if there are moving objects in the room or if the speaker or
microphone is moving. This convolution is followed by a non-linear
distortion in the microphone and its preamplifier and convolved again with
another impulse response, this time representing a frequency and phase
response of recorder's amplifiers. Finally it is non-linearly distorted in
the process of making a magnetic recording on the tape. Another batch of
distortions comes during the playback. On top of that, in every step a certain
amount of additive noise is added. Clearly, modeling, let alone reverting these
interferences is a difficult task.

@@

In order to perform clustering on any dataset, a metric on the data is needed.
We have used that proposed by Mandel \& Ellis\cite{mandel2005song}, which
is based on mel-frequency cepstra. A distance matrix has been created using the
tool musly\cite{schnitzer2011using} and the clusters on top of that using
hierarchical clustering\cite{johnson1967hierarchical}.

@@

The small but apparent reduction in word error rate in case of the overdriven
cluster is the positive, promising result. The improvement is rather small, and
as such it does not seem worthwhile applying it on the entire corpus. It
proves, however, that there is some potential. An important immediate benefit is
that the transferred overdriven (or should we say de-overdriven) sound files are
much easier on the ear and it can be a help for human listeners.

The wrecking word error rate increase from 69\% to 94\% in case of the
recordings taken at low tape speed requires further investigation.
Figure~\ref{fig:plzen} shows the waveform and spectrum of a sample from the low-speed
cluster before and after the transfer. For comparison,
Figure~\ref{fig:overdrive} shows the same for the overdriven cluster.
Notice how parts of the signal are completely missing in the converted low-speed
sample. It seems that when the signal is too hard to separate from the noise,
the covertor prefers to generate silence.

@@

Neural-network domain transfer offers a universally applicable method to
mitigate any kind of and combination of acoustic defects without the necessity
of modeling the specific effects.
The method was demonstrated to help to a small extent with speech recognition
results. However, where the damage is too severe, it discards the only
information at hand instead of improving its quality.

@@

@@

This work has been using language resources developed, stored and distributed by
the LINDAT/CLARIAH project of the Ministry of Education, Youth and Sports of the
Czech Republic (project LM2018101).

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Speech Enhancement for Low Quality Audio}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Rincewind\inst{1}\orcidID{0000-0000-0000-0000}
\and Luggage\inst{1}\orcidID{0000-0000-0000-0001}}
%
\authorrunning{Rincewind \and Luggage}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Unseen University,\\
Ankh-Morpork,\\
The Discworld\\
\email{\{rincewind,luggage\}@unseen-university.ankh-morpork.discworld}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The paper presents an attempt to solve the problem of bad audio quality
affecting the listening experience and accuracy of speech-to-text systems. We
describe two experiments: one with automated spectral subtraction and one
with Gycle-GAN domain transfer from the defect to the
pure as a way to reduce the impact of noise, overdrive, reverberation and other
defects.
\keywords{speech recognition, GycleGAN, domain transfer, signal processing}
\end{abstract}

@@

There are in principle two ways to deal with the negative effect of acoustic
deficiencies in speech recognition: 1) to adapt the model, which includes
training on augmented data, and 2) to adapt the input data. This article
presents one experiment in the former and one in the latter approach.

@@

The application is at the spoken corpus of Karel Makoň\cite{makondata}, for which
a speech recognition system has been developed\cite{kruuza2012making} but where
the heterogenous acoustic quality affects both the speech recognition
performance and the listening experience. Whereas the word error rate on the
overall test set is 19\%, it is 45\% on a sample of recordings suffering from
overdrive, and 68\% on recordings taken on a magnetophone tape with a slow
recording speed of 2 centimeters per second after some 30 years of living-room
storage.

@@

The individual interference types affect each other and can occur several times
in several stages. For instance, the echo is essentially convolution with a certain vector of
amplitude values, known as the impulse response. It might even not be
constant if there are moving objects in the room or if the speaker or
microphone is moving. This convolution is followed by a non-linear
distortion in the microphone and its preamplifier and convolved again with
another impulse response, this time representing a frequency and phase
response of recorder's amplifiers. Finally it is non-linearly distorted in
the process of making a magnetic recording on the tape. Another batch of
distortions comes during the playback. On top of that, in every step a certain
amount of additive noise is introduced. Clearly, modeling, let alone reverting these
interferences is a difficult task.

\section{Denoising}

We conducted a baseline experiment employing the standard denoising based on
spectral subtraction. We assumed that each recording in the
corpus suffers from a consistent kind of noise. This does not hold for all the
recordings but we dismissed this fact for the sake of simplicity. We used sox
fox the denoising routine. The process for each recording is as follows:
\begin{enumerate}
\item{Identify and isolate a sample of pure stationary noise,}
\item{extract the noise profile,}
\item{apply noise reduction based on the noise profile.}
\end{enumerate}

Items 2 and 3 are taken care of by sox. As for extracting fitting silences, we
used the following method:
\begin{enumerate}
\item{
    Identify and extract all precicted silences using an
    existing aligned automatic transcript.
}
\item{
    Select 100 silences around the 25th percentile ordered by length.
    This secures silences that are neither too short nor too long. Long silences
    tend to contain non-stationary noises, that's why we avoid using them.
}
\item{
    Create a distance matrix on the silences using
    musly\cite{schnitzer2011using}.
}
\item{
    Select 10 silences with the lowest median on the distance matrix. These
    silences are most like other silences, thus being least likely to contain
    non-speech events.
}
\item{
    Finally, catenate and output the selected silences.
}
\end{enumerate}

Subjective evaluation confirms the expectable outcome that relatively
high-quality recordings that only suffer from some additive noise get
easier to listen to. Recordings suffering from other defects and overall lower
audio quality sometimes get rather harder to understand.

We have also trained a speech recognition system on the denoised data. The
performance has sunken to word error rate of 0.94, which basically means the
transcription doesn't work at all. Probably with some parameter tweaking this
could be remedied but it doesn't seem to be a promising path. Curiously enough, the
original model trained on non-denoised data performs much better on the denoised
test data with word error rate of 0.70. Table~\ref{tab:results-denoise} sums
these up.

\begin{table}[htpb]
\begin{center}
\begin{tabular}{|l||r|r|}
\hline
WER    & original model & denoised model \\
\hline
original testing data & 0.189 & 0.940 \\
denoised testing data & 0.698 & 0.941 \\
\hline
\end{tabular}
\caption{Word error rate for the original and denoised test data by a model trained
on the original and denoised training data.}\label{tab:results-denoise}
\end{center}
\end{table}

@@

In order to perform clustering on any dataset, a metric on the data is needed.
We have used that proposed by Mandel \& Ellis\cite{mandel2005song}, which
is based on mel-frequency cepstra. A distance matrix has been created using
musly and the clusters on top of that using
hierarchical clustering\cite{johnson1967hierarchical}.

@@

The word error rate decrease in the overdriven cluster is unfortunately statistically
insignificant with standard deviations 0.054 and 0.079 of the transferred and
original versions respectively. An important benefit is
that the transferred overdriven (or should we say de-overdriven) sound files are
much easier on the ear and it can be a help for human listeners.

The wrecking word error rate increase from 69\% to 94\% in case of the
recordings taken at low tape speed requires further investigation.
Figure~\ref{fig:plzen} shows the waveform and spectrum of a sample from the low-speed
cluster before and after the transfer. For comparison,
Figure~\ref{fig:overdrive} shows the same for the overdriven cluster.
Notice how parts of the signal are completely missing in the converted low-speed
sample. It seems that when the signal is too hard to separate from the noise,
the convertor prefers to generate silence.

@@

Neural-network domain transfer offers a universally applicable method to
mitigate any kind and combination of acoustic defects without the necessity
of specifically modeling them. However,
we have not recorded any significant increase in speech recognition performance.
Also, where the damage is too severe, the processing discards the only
information at hand instead of improving its quality.

@@

The conducted experiments on mitigating defects in acoustic quality of recorded speech
show that 1) it is not easy to reach an increase in speech recognition
performance, 2) it is easier to reach an improvement for the subjective
human listening experience and 3) heavily impaired recordings are particularly difficult to
make any better.

@@

This work has been supported by undisclosed institutions and entities.
